Measuring individual differences in the understanding of gaze cues

### 2022-08-17 paper/manuscript.Rmd
first complete draft of discussion

---
title             : "Measuring individual differences in the understanding of gaze cues across the lifespan"
shorttitle        : "gaze understanding"

author: 
  - name          : "Julia Prein"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany"
    email         : "julia_prein@eva.mpg.de"

  - name          : "Manuel Bohn"
    affiliation   : "1"

  - name          : "Steven Kalinke"
    affiliation   : "1"

  - name          : "Daniel M. Haun"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"

authornote: |

abstract: |
  To explain and predict the behavior of agents, we use social cognition: we represent and reason about others' perspectives, knowledge, intentions, beliefs, and preferences. However, traditional measures of social cognition (e.g., false belief change-of-location tasks) often lack satisfactory psychometric properties: they are not designed to capture variation *between* children and rely on low trial numbers, dichotomous measures, and group averages. This has profound implications for what these studies can show. Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change. To fully understand how social-cognitive abilities emerge and relate to each other, we need new tools that can reliably measure individual differences.
   We designed a gaze understanding task to study social cognition in young children and adults to approach this issue. We concentrate on an essential ability that is involved in many social-cognitive reasoning processes: gaze understanding – the ability to locate and use the attentional focus of an agent. Our interactive task works across devices and enables supervised and unsupervised, as well as in-person and remote testing. The implemented spatial layout allows for discrete and continuous measures of participants’ click imprecision and is easily adaptable to different study requirements. 
  Here we show that our task induces inter-individual differences in a child (N = XXX) and an adult (N = XXX) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test-retest reliability estimates underline that the captured variation is systematic. This work shows a promising way forward in the study of individual differences in social cognition and will help us explore the in(ter)dependence of our core social-cognitive processes in greater detail.

keywords          : "social cognition, individual differences, gaze cues, psychometrics"
wordcount         : "X"

bibliography      : ["../../references.bib"]

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf

header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
---

@@

# Introduction 
Social cognition -- representing and reasoning about an agent’s perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior -- is among the most-studied phenomena in developmental research. 
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development [@gopnik1991young; @wellman2001metaanalysis; @wellman2004scaling; @peterson2012mind; @rakoczy2022foundations].
Yet, there are always individual differences. 
Identifying variability in social-cognitive abilities and factors influencing their development is vital in theory building (e.g., to test causal predictions) and for designing interventions 
[@happe2017structure; @mundy2007individual; @kidd2018individual; @underwood1975individual; @lecce2014promoting].

@@

However, frequently, developmental psychologists are surprised to find minor or no association between measures of social cognition that are thought to be theoretically related -- cross-sectionally and/or longitudinally [e.g., @sodian2016understanding]. 
This might be because traditional measures of social cognition are not designed to capture variation between children: 
they often rely on low trial numbers, small sample sizes, and dichotomous measures. A recent review showed that many atudies on social cognition measures failed to report relevant psychometric properties at all [@beaudoin2020systematic] or -- when they did -- showed mixed results on test-retest reliability [@mayes1996testretest; @hughes2000good].

To give an example: perhaps the most commonly applied prototypical measure for social cognition is the change-of-location false belief task [@baron-cohen1985does; @wimmer1983beliefs]. 
Here, children watch a short sequence of events (often acted out or narrated by the experimenters). 
A doll called Sally puts her marble into a basket. 
After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box. 
Participants then get asked where Sally will look for her marble once she returns. 
The outcome measures false belief understanding in a dichotomous way: 
children pass the task if they take the protagonist’s epistemic state into account and answer that she will look into the basket. 
Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age [for a review, see @wellman2001metaanalysis]. 
Several cross-cultural studies supported this evidence [@barrett2013early; @callaghan2005synchrony; cf. @mayer2015weird].

However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value [@repacholi2003introduction]. 
Thus, this sort of task seems to be well suited to track group-level developmental trends, yet it fails to capture individual differences [cf. "reliability paradox", @hedge2018reliability]. 
As @wellman2012theory put it, “it’s really only passing/failing one sort of understanding averaged across age” (p. 317). 
This has profound implications for what studies on individual differences using this task (or others) can show. 
Poor measurement of social cognition on an individual level is likely to conceal relations between different aspects of cognition and may obscure developmental change.

@@

@schaafsma2015deconstructing compiled a "wish-list" for new social-cognitive paradigms. 
They advocated for parametric -- instead of dichotomous -- measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates [see also @beaudoin2020systematic; @hughes2015individual]. 
New tasks should induce variation across age groups, including older children and adults [@repacholi2003introduction]. 
Another goal in creating new tasks should be to focus on "face value": 
measures should probe the underlying social-cognitive ability as straight-forward and directly as possible. 
The task should serve as a proxy for behavior as it appears in the real world and should be validated in relation to real-world experiences [@repacholi2003introduction].

# A new measure of gaze understanding

Our goal was to design a new measure of social cognition that captures individual differences across age groups in a systematic and reliable way. 
We focused on one of the fundamental abilities implicated in many social-cognitive reasoning processes: 
gaze understanding -- the ability to locate and use the attentional focus of an agent. 
The first component of this ability is often termed gaze following -- turning one's eyes in the same direction as the gaze of another agent -- 
and has been studied intensively [@itakura1998use; @delbianco2019developmental; @byers-heinlein2021development; @moore2008development; @coelho2006searching; @tomasello2007reliance; @hernik2019infant; @astor2021gaze; @frischen2007gaze; @shepherd2010following; @lee1998children]. 
Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a "front end ability" [@brooks2005development, p. 535]. 
In our definition, gaze understanding goes one step further by including the *acting on the gaze-cued location* -- therefore, using the available social information to guide one's behavior as needed in real-life conditions.

@@

# Task design

## Implementation
The code is open-source (https://github.com/ccp-eva/gafo-demo) and a live demo version can be found under: https://ccp-odc.eva.mpg.de/gafo-demo/.

@@

## Stimuli
Our newly implemented task features an online game where children or adults are asked to search for a balloon. The events proceed as follows (see Figure \@ref(fig:fig1)B and C). An animated agent (a sheep, monkey, or pig) looks out of a window of a house. A balloon (i.e., target; blue, green, yellow, or red) is located in front of them. The target then falls to the ground. At all times, the agent’s gaze tracks the movement of the target. That is, the pupils and iris of the agent move in a way that their center aligns with the center of the target. While the distance of the target's flight depends on the final location, the target moves at a constant speed. Participants are then asked to locate the target: they respond by touching or clicking on the screen. Visual access to the target’s true location is manipulated by a hedge. Participants either have full, partial, or no visual access to the true target location. When partial or no information about the target location is accessible, participants are expected to use the agent’s gaze as a cue.

To keep participants engaged and interested, the presentation of events is accompanied by cartoon-like effects. Each trial starts with an attention-getter: an eye-blinking sound plays while the pupils and iris of the agent enlarge (increase to 130%) and change in opacity (decrease to 75%) for 0.3 sec. The landing of the target is accompanied by a tapping sound. Once the target landed, the instructor's voice asks "Where is the balloon?". For confirming the participant's click, a short plop sound plays and a small orange circle appears at the location of choice. If no response is registered within 5 secs after the target landed, an audio prompt reminds the participant to respond.

## Trials
Trials differ in the amount of visual access that participants have to the final target position. Before the test trials start, participants complete four training trials during which they familiarize themselves with clicking the screen. In the first training trial, participants have full visual access to the target flight and the target's end location and are simply asked to click on the visible balloon. In the second and third training trials, participants have partial access: they witness the target flight but cannot see the target's end location. They are then asked to click on the hidden balloon, i.e., the location where they saw the target land. In test trials, participants have no visual access to the target flight or the end location. Participants are expected to use the agent’s gaze as a cue to locate the target. The first trial of each type comprises a voice-over description of the presented events. The audio descriptions explicitly state that the agent is always looking at the target (see Appendix for audio script). After the four training trials, participants receive 15 test trials. The complete sequence of four training trials and 15 test trials can be easily completed within 5-10 minutes.

## Study versions
We designed two study versions which differ in the final hiding place of the target and, consequently, on the outcome measure: a *hedge version* (continuous) and a *box version* (discrete). Both versions use the same first training trial and then differ in the consecutive training and test trials. 
In the hedge version, participants have to indicate their estimated target location directly on a hedge. Here, the dependent variable is imprecision, which is defined as the absolute difference between the target center and the x coordinate of the participant’s click.
In the box version, the target lands in a box and participants are asked to click on the  box that hides the target. Researchers have the choice how many boxes are shown: one up to eight boxes can be displayed as potential hiding locations. Here, we use a categorical outcome (i.e., which box was clicked) to calculate the proportion of correct responses. 
Note that in the test trials of both versions, the target flight is covered by a hedge. In the hedge version, the hedge then shrinks to a minimum height required to cover the target's end location. In the box version, the hedge shrinks completely. The boxes then hide the target's final destination (see Figure \@ref(fig:fig1)B and C).

(ref:figlab1) **Study setup**. 
(A) Infrastructure for online testing. (i) Subjects aged 3 – 99+ can participate. Data collection can take place anywhere: online, in kindergartens or research labs. (ii) The task is presented as a website that works across devices. (iii) The scripts for the website and the recorded data are stored on secure local servers.
(B) Hedge version (continuous) of the gaze understanding task. (i) The agent stands in a window with the target in front of them. (ii) A hedge grows and covers the target. (iii) The target falls to a random location on the ground. The agent's eyes track the movement of the target. 
(C) Box version (discrete) of the gaze understanding task. Number of boxes (min. 1; max. 8) as potential hiding locations can be set according to the researcher's need.

@@

# Individual differences
Our first aim was to assess whether our gaze understanding task induces inter-individual variation in a child and adult sample. Furthermore, we were interested in how the data collection mode influences responses.

Methods, sample size and analysis were pre-registered: https://osf.io/snju6 (child sample) and https://osf.io/r3bhn (adult sample). Participants were equally distributed across the two study versions. The study was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology. Data was collected between May and October 2021.

@@

vali_adults <- vali %>%
  filter(ageInYears == 18) %>% 
  mutate(
    gender = ifelse(is.na(gender), 0, paste(gender)), 
    female = ifelse(gender == "f", 1, 0),
  ) %>% 
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(age, na.rm = T) %>% round(2),
    sdAge = sd(age, na.rm = T) %>% round(2),
    minAge = min(age, na.rm = T),
    maxAge = max(age, na.rm = T),
    nFemales = ceiling(sum(female/15)),
  )
```
## Participants
We collected data from an in-person child sample, a remote child sample, and a remote adult sample.
In-person testing with children took place in kindergartens in Leipzig, Germany. The in-person child sample consisted of 
`r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised"])` children, including 
`r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` 3-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` girls),
`r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` 4-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` girls),
and `r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` 5-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` girls).

For our remote child sample, we recruited families via an internal database. The remote child sample included `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised"])` children, including 
`r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` 3-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` girls),
`r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` 4-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` girls),
and `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` 5-year-olds 
(mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` months, 
SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])`, 
range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])`
- `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])`, 
`r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` girls). Children in our sample grow up in an industrialized, urban Central-European context. Information on socioeconomic status was not formally recorded, although the majority of families come from mixed, mainly mid to high socioeconomic backgrounds with high levels of parental education.

>>>>

---
title             : "Measuring individual differences in the understanding of gaze cues across the lifespan"
shorttitle        : "gaze understanding"
author:

  - name          : "Julia Prein"
    affiliation   : "1"
    corresponding : yes    # Define only one corresponding author
    address       : "Max Planck Institute for Evolutionary Anthropology, Deutscher Platz 6, 04103 Leipzig, Germany"
    email         : "julia_prein@eva.mpg.de"

  - name          : "Manuel Bohn"
    affiliation   : "1"
  - name          : "Steven Kalinke"
    affiliation   : "1"

  - name          : "Daniel M. Haun"
    affiliation   : "1"
affiliation:
  - id            : "1"
    institution   : "Department of Comparative Cultural Psychology, Max Planck Institute for Evolutionary Anthropology, Leipzig, Germany"
authornote: |
abstract: |

  To explain and predict the behavior of agents, we use social cognition: we represent and reason about others' perspectives, knowledge, intentions, beliefs, and preferences. However, traditional measures of social cognition (e.g., false belief change-of-location tasks) often lack satisfactory psychometric properties: they are not designed to capture variation *between* children and rely on low trial numbers, dichotomous measures, and group averages. This has profound implications for what these studies can show. Poor measurement of social cognition on an individual level may conceal relations between different aspects of cognition and may obscure developmental change. To fully understand how social-cognitive abilities emerge and relate to each other, we need new tools that can reliably measure individual differences.
   We designed a gaze understanding task to study social cognition in young children and adults to approach this issue. We concentrate on an essential ability that is involved in many social-cognitive reasoning processes: gaze understanding – the ability to locate and use the attentional focus of an agent. Our interactive task works across devices and enables supervised and unsupervised, as well as in-person and remote testing. The implemented spatial layout allows for discrete and continuous measures of participants’ click imprecision and is easily adaptable to different study requirements. 
  Here we show that our task induces inter-individual differences in a child (N = XXX) and an adult (N = XXX) sample. Our two study versions and data collection modes yield comparable results that show substantial developmental gains: the older children are, the more accurately they locate the target. High internal consistency and test-retest reliability estimates underline that the captured variation is systematic. This work shows a promising way forward in the study of individual differences in social cognition and will help us explore the in(ter)dependence of our core social-cognitive processes in greater detail.
keywords          : "social cognition, individual differences, gaze cues, cognitive development"
wordcount         : "X"
bibliography      : ["../../references.bib"]
floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no
documentclass     : "apa6"
classoption       : "man"
output            : papaja::apa6_pdf
header-includes:
  - \usepackage{setspace}
  - \captionsetup[figure]{font={stretch=1}}
editor_options: 
  markdown: 
    wrap: sentence
---

@@

# Introduction
Social cognition -- representing and reasoning about an agent's perspectives, knowledge states, intentions, beliefs, and preferences to explain and predict their behavior -- is among the most-studied phenomena in developmental research.
In recent decades, much progress has been made in determining the average age at which a specific social-cognitive ability emerges in development [@gopnik1991young; @wellman2001metaanalysis; @wellman2004scaling; @peterson2012mind; @rakoczy2022foundations]. Yet, there are always individual differences. Identifying variability in social-cognitive abilities and factors influencing their development is vital in theory building (e.g., to test causal predictions) and for designing interventions [@happe2017structure; @mundy2007individual; @kidd2018individual; @underwood1975individual; @lecce2014promoting].

@@

However, frequently, developmental psychologists are surprised to find minor or no association between measures of social cognition that are thought to be theoretically related -- cross-sectionally and/or longitudinally [e.g., @sodian2016understanding]. This might be because traditional measures of social cognition are not designed to capture variation between children: they often rely on low trial numbers, small sample sizes, and dichotomous measures. A recent review showed that many studies on social cognition measures failed to report relevant psychometric properties at all [@beaudoin2020systematic] or -- when they did -- showed mixed results on test-retest reliability [@mayes1996testretest; @hughes2000good].

To give an example: perhaps the most commonly applied prototypical measure for social cognition is the change-of-location false belief task [@baron-cohen1985does; @wimmer1983beliefs]. Here, children watch a short sequence of events (often acted out or narrated by the experimenters). A doll called Sally puts her marble into a basket. After Sally leaves the scene, a second doll named Anne takes the marble and moves it into a box. Participants then get asked where Sally will look for her marble once she returns. The outcome measures false belief understanding in a dichotomous way: children pass the task if they take the protagonist's epistemic state into account and answer that she will look into the basket. Many years of research utilizing these verbal change-of-location tasks suggest that children develop belief-representing abilities at four to five years of age [for a review, see @wellman2001metaanalysis]. Several cross-cultural studies supported this evidence [@barrett2013early; @callaghan2005synchrony; cf. @mayer2015weird].

However, from this age onwards, the change-of-location task shows ceiling effects and has very limited diagnostic value [@repacholi2003introduction]. Thus, this sort of task seems to be well suited to track group-level developmental trends, yet it fails to capture individual differences [cf. "reliability paradox", @hedge2018reliability].
As @wellman2012theory put it, "it's really only passing/failing one sort of understanding averaged across age" (p. 317).
This has profound implications for what studies on individual differences using this task (or others) can show.
Poor measurement of social cognition on an individual level is likely to conceal relations between different aspects of cognition and may obscure developmental change.

@@

@schaafsma2015deconstructing compiled a "wish-list" for new social-cognitive paradigms.
They advocated for parametric -- instead of dichotomous -- measures covering proficiency as a range, avoiding floor and ceiling effects, and showing satisfactory test-retest reliability estimates [see also @beaudoin2020systematic; @hughes2015individual].
New tasks should induce variation across age groups, including older children and adults [@repacholi2003introduction].
Another goal in creating new tasks should be to focus on the "face value": measures should probe the underlying social-cognitive ability as straight-forward and directly as possible.
Keeping task demands minimal is also beneficial for using the paradigm in a variety of different cultural, clinical and demographic contexts [@molleman2019unleashing].
The task should serve as a proxy for behavior as it appears in the real world and should be validated in relation to real-world experiences [@repacholi2003introduction].

# A new measure of gaze understanding
Our goal was to design a new measure of social cognition that captures individual differences across age groups in a systematic and reliable way.
We focused on a fundamental ability implicated in many social-cognitive reasoning processes: gaze understanding -- the ability to locate and use the attentional focus of an agent.
The first component of this ability is often termed gaze following -- turning one's eyes in the same direction as the gaze of another agent -- and has been studied intensively [@itakura1998use; @delbianco2019developmental; @byers-heinlein2021development; @moore2008development; @coelho2006searching; @tomasello2007reliance; @hernik2019infant; @astor2021gaze; @frischen2007gaze; @shepherd2010following; @lee1998children].
Following an agent's gaze provides insights into their intentions, thoughts, and feelings by acting as a "front end ability" [@brooks2005development, p. 535].
In our definition, gaze understanding goes one step further by including the *acting on the gaze-cued location* -- therefore, using the available social information to guide one's behavior as needed in real-life conditions.

@@

# Task design
## Implementation
The code is open-source (<https://github.com/ccp-eva/gafo-demo>) and a live demo version can be found under: <https://ccp-odc.eva.mpg.de/gafo-demo/>.

@@

## Stimuli
Our newly implemented task features an online game where children or adults are asked to search for a balloon.
The events proceed as follows (see Figure \@ref(fig:fig1)B and C).
An animated agent (a sheep, monkey, or pig) looks out of a window of a house.
A balloon (i.e., target; blue, green, yellow, or red) is located in front of them.
The target then falls to the ground.
At all times, the agent's gaze tracks the movement of the target.
That is, the pupils and iris of the agent move in a way that their center aligns with the center of the target.
While the distance of the target's flight depends on the final location, the target moves at a constant speed.
Participants are then asked to locate the target: they respond by touching or clicking on the screen.
Visual access to the target's true location is manipulated by a hedge.
Participants either have full, partial, or no visual access to the true target location.
When partial or no information about the target location is accessible, participants are expected to use the agent's gaze as a cue.

To keep participants engaged and interested, the presentation of events is accompanied by cartoon-like effects.
Each trial starts with an attention-getter: an eye-blinking sound plays while the pupils and iris of the agent enlarge (increase to 130%) and change in opacity (decrease to 75%) for 0.3 sec.
The landing of the target is accompanied by a tapping sound.
Once the target landed, the instructor's voice asks "Where is the balloon?".
For confirming the participant's click, a short plop sound plays and a small orange circle appears at the location of choice.
Participants do not receive differential feedback so that learning effects are reduced and trials stay comparable across the sample.
If no response is registered within 5 secs after the target landed, an audio prompt reminds the participant to respond.

## Trials
Trials differ in the amount of visual access that participants have to the final target position.
Before the test trials start, participants complete four training trials during which they familiarize themselves with clicking the screen.
In the first training trial, participants have full visual access to the target flight and the target's end location and are simply asked to click on the visible balloon.
In the second and third training trials, participants have partial access: they witness the target flight but cannot see the target's end location.
They are then asked to click on the hidden balloon, i.e., the location where they saw the target land.
In test trials, participants have no visual access to the target flight or the end location.
Participants are expected to use the agent's gaze as a cue to locate the target.
The first trial of each type comprises a voice-over description of the presented events.
The audio descriptions explicitly state that the agent is always looking at the target (see Appendix for audio script).
After the four training trials, participants receive 15 test trials.
The complete sequence of four training trials and 15 test trials can be easily completed within 5-10 minutes.

## Study versions
We designed two study versions which differ in the final hiding place of the target and, consequently, on the outcome measure: a *hedge version* (continuous) and a *box version* (discrete).
Both versions use the same first training trial and then differ in the consecutive training and test trials.
In the hedge version, participants have to indicate their estimated target location directly on a hedge.
Here, the dependent variable is imprecision, which is defined as the absolute difference between the target center and the x coordinate of the participant's click.
In the box version, the target lands in a box and participants are asked to click on the box that hides the target.
Researchers have the choice how many boxes are shown: one up to eight boxes can be displayed as potential hiding locations.
Here, we use a categorical outcome (i.e., which box was clicked) to calculate the proportion of correct responses.
Note that in the test trials of both versions, the target flight is covered by a hedge.
In the hedge version, the hedge then shrinks to a minimum height required to cover the target's end location.
In the box version, the hedge shrinks completely.
The boxes then hide the target's final destination (see Figure \@ref(fig:fig1)B and C).

(ref:figlab1) **Study setup**. (A) Infrastructure for online testing. (i) Subjects aged 3 -- 99+ can participate. Data collection can take place anywhere: online, in kindergartens or research labs. (ii) The task is presented as a website that works across devices. (iii) The scripts for the website and the recorded data are stored on secure local servers. (B) Hedge version (continuous) of the gaze understanding task. (i) The agent stands in a window with the target in front of them. (ii) A hedge grows and covers the target. (iii) The target falls to a random location on the ground. The agent's eyes track the movement of the target. (C) Box version (discrete) of the gaze understanding task. Number of boxes (min. 1; max. 8) as potential hiding locations can be set according to the researcher's need.

@@

# Individual differences
Our first aim was to assess whether our gaze understanding task induces inter-individual variation in a child and adult sample.
Furthermore, we were interested in how the data collection mode (in-person vs. remote) influences responses.

Methods, sample size and analysis were pre-registered: <https://osf.io/snju6> (child sample) and <https://osf.io/r3bhn> (adult sample).
Participants were equally distributed across the two study versions.
The study was approved by an internal ethics committee at the Max Planck Institute for Evolutionary Anthropology.
Data was collected between May and October 2021.

@@

vali_adults <- vali %>%
  filter(ageInYears == 18) %>% 
  ungroup() %>% 
  mutate(
    gender = ifelse(is.na(gender), 0, paste(gender)), 
    female = ifelse(gender == "f", 1, 0),
  ) %>% 
  summarise(
    nTotal = n_distinct(subjID), 
    meanAge = mean(age, na.rm = T) %>% round(2),
    sdAge = sd(age, na.rm = T) %>% round(2),
    minAge = min(age, na.rm = T),
    maxAge = max(age, na.rm = T),
    nFemales = ceiling(sum(female/15)),
  )
```

## Participants
We collected data from an in-person child sample, a remote child sample, and a remote adult sample.
In-person testing with children took place in kindergartens in Leipzig, Germany.
The in-person child sample consisted of `r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised"])` children, including `r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` 3-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "3"])` girls), `r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` 4-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "4"])` girls), and `r sum(vali_kids$nTotal[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` 5-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "in-person - supervised" & vali_kids$ageInYears == "5"])` girls).

For our remote child sample, we recruited families via an internal database of children living in Leipzig, Germany, whose parents volunteered to participate in studies on child development and who indicated interest in online studies.
Families received an email with a short study description and a personalized link, and received a reminder two weeks later if they had not already participated in the study.

The remote child sample included `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised"])` children, including `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` 3-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "3"])` girls), `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` 4-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "4"])` girls), and `r sum(vali_kids$nTotal[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` 5-year-olds (mean = `r sum(vali_kids$meanAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` months, SD = `r sum(vali_kids$sdAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])`, range = `r sum(vali_kids$minAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` - `r sum(vali_kids$maxAgeInMonths[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])`, `r sum(vali_kids$nFemales[vali_kids$datacollection == "remote - unsupervised" & vali_kids$ageInYears == "5"])` girls).
Children in our sample grow up in an industrialized, urban Central-European context.
Information on socioeconomic status was not formally recorded, although the majority of families come from mixed, mainly mid to high socioeconomic backgrounds with high levels of parental education.
